Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread

### 2018-02-09 paper.tex
typeset error fixed

\gmj{I had saved these notes from an earlier version of the paper.  There is plenty of evidence of hybridization and introgression (both farmer-mediated and and natural), and some adaptive traits have been conferred, but I could not find direct evidence of specific cases of adaptive natural introgression from wild relatives to cultivated potatoes. This new paper fills that gap, if we decide that the paper's merits are substantial enough that we can overlook its flaws.}
\gmj{If we decide to keep and use the new potato introgression publication, that would be the natural choice for the potato reference, whether or not we include the write-up of the potato case study.}

Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000 BP, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}.
The northern members of the polyphyletic \emph{S. brevicaule} wild potato complex have been identified as likely progenitors of modern potatoes \cite{correll1962potato}, but determination of a single progenitor species is unlikely, either because widespread gene flow in the complex will mask signs of this phylogeny or because \emph{s. tuberosum} has a polyphyletic origin.
Also, these northern members of the \emph{S. breviaule} complex are not clearly defined, and may in fact be one singular species (in which case the species name would be \emph{S. bukasovii}, \cite{spooner2005single}).

@@

Introgressive hybridization is widespread in potatoes \cite{grun1990evolution}.
The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}.
Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}.
Cultivated diploid potatoes have been shown introgressed with genes from \emph{S. sparsipilum} \cite{rabinowitz1990high} and \emph{S. megistacrolobum} \cite{johns1987relationships, huaman1980solanum}.
Drought- and frost-resistance genes introgressed from wild relatives expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}.
Resistance genes have been experimentally introgressed from wild poptato relatives (\emph{S. tuberosum subsp. andigena}, \cite{van1999tight}, \emph{S. bulbocastanum}, \cite{van2003ancient}) into cultivated potato, although this process is hindered by compatability issues including ploidy levels and Endosperm Balance Number \cite{johnston1980significance}.

>>>>

Modern potato (\emph{Solanum tuberosum}) is believed to have been domesticated in southern Peru in sympatry with a multitude of wild relatives about 6000-10000 BP, although the exact location and formal classification and phylogenetic relationships between these taxa have long been disputed \cite{huaman2002reclassification, spooner2005single, pickersgill1977origins, hawkes1988evolution}.
A distance-based phylogeny assembled from a SNP array from a \emph{Solanum} diversity panel (which included wild, landrace, and cultivated species) formed three groups (North American, South American, and cultivated) and identified \emph{S. candolleanum} as being most probable wild progenitor of the potato \cite{hardigan2015taxonomy}.

@@

Nevertheless, introgressive hybridization is widespread in potatoes \cite{grun1990evolution}.
The various cultivars of Andean potatoes are interfertile, forming one large plastic gene pool \cite{quiros1992increase}.
%Cultivated diploid potatoes have been shown introgressed with genes from \emph{S. sparsipilum} \cite{rabinowitz1990high} and \emph{S. megistacrolobum} \cite{johns1987relationships, huaman1980solanum}.
%Resistance genes have been experimentally introgressed from wild poptato relatives (\emph{S. tuberosum subsp. andigena}, \cite{van1999tight}, \emph{S. bulbocastanum}, \cite{van2003ancient}) into cultivated potato, although this process is hindered by compatability issues including ploidy levels and Endosperm Balance Number \cite{johnston1980significance}.
%Hardigan and colleagues \cite{hardigan2017genome} find evidence of extensive adaptive introgression into cultivated potatoes.
In their wild/landrace/domesticated diversity panel, Hardigan and colleagues \cite{hardigan2017genome} found a high level of genomic contribution from wild species (including \emph{S. candolleanum}) into Andean tetraploid cultivated landraces and North American cultivars (20.2\% and 32.1\%, respectively).
Genes located within these introgressed regions were more likely to be highly-expressed, stress-inducible, and contained loci related to disease resistance, drought tolerance, and heat tolerance, suggesting that these introgressed regions impart stress response mechanisms critical to survival, possibly facilitating tolerance for new environmental pressures during range expansion.
Drought- and frost-resistance genes introgressed from wild relatives are also reported to have expediated the spread of cultivated potatoes into the central Andean Altiplano \cite{johns1986ongoing, hawkes1962origin, schmiediche1980breeding}.
Andean potatoes exhibit high ecological versatility, due in part either to alleleic diversity in polyploids or introgression of desirable alleles from wild relatives in diploids \cite{zimmerer1998ecogeography}.
